---
title: "Simulations"
author: "Andrew McDavid"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
---

# Scenarios
## Graphs
We simulate from:

1.  Chain graphs of varying dimensions
2.  A subgraph of reported gene-gene interactions from E Coli

## Generative distributions
We consider subclasses of the graphical hurdle model, and moderate departures from the class.

1.  Full hurdle model
2.  Logistic only
3.  (1) contaminated with a scaled $t_8$ distribution
4.  A synthetic "1000" cell dataset, comprised of models 1-2

## Estimators
1.  Graphical hurdle
2.  Neighborhood selection with $\ell_1$-penalized logistic regression
3.  Neighborhood selection with $\ell_1$-penalized linear regression
4.  Neighborhood selection with $\ell_1$-penalized linear regression using a Gaussian Copula (non-paranormal)
5.  Mutual information-based (ARACNE etc, use netbenchmark)

## Evaluation
1. False positives/negatives with oracle tuning
2. False positives/negatives with BIC-based tuning

```{r library}
library(HurdleNormal)
source('simulation_library.R')
```

```{r ecoli}
set.seed(5678)
vertex_seq <- sample(1565, size=500)
options(mc.cores=6)
gnw <- ecoli_genenetweaver(vertex_seq=vertex_seq)
```


